---
title: "PTB Group 9.3 Staff Statisic"
author: "Arne Diedrich"
date: "Version Date `r Sys.Date()` - Data: 2024-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggpubr)
library(stargazer)
library(knitr)
library(rcompanion)

##### Importing Data ##### 
personal <- read_excel("data/Personal.xlsx", sheet = "Personal")
metadata <- read_excel("data/Personal.xlsx", sheet = "Metadaten")

##### Organizing Data ####
personal$Position = as.factor(personal$Position) 
personal$Position = factor(personal$Position, levels = c("RL", "FB", "PK", "PA", "Head of Group 9.3", "Knowledge Manager", "Assistent"))

##### Figs ####
#By Section
fig_bySection = personal %>% group_by(Section) %>% ggplot(aes(Section, fill = Sex)) + geom_bar() + labs(y=NULL,title = "Fig. 1 - Gender Distribution by Section") + scale_y_continuous(n.breaks = 10) + theme(legend.position = "none")

#personal %>% group_by(Section) %>% ggplot(aes(Section, fill = Sex)) + geom_bar(position = position_fill())

#By Position (Without 9.3)
fig_position_bySection = personal %>% filter("9.3" != Section) %>% group_by(Position) %>% ggplot(aes(Position, fill = Sex)) + 
  geom_bar() + facet_wrap(.~ Section) + labs(y = NULL, x = NULL)

fig_position_gesamt = personal %>% filter("9.3" != Section) %>% group_by(Position) %>% ggplot(aes(Position, fill = Sex)) + 
  geom_bar() + labs(y = NULL, title = "Fig. 2 - Gender Distribution by Position (only Section Staff)") + scale_y_continuous(breaks = seq(0,38,2)) + 
  scale_x_discrete(labels = metadata$Long)


ggarrange(fig_position_gesamt, fig_position_bySection)
```

This Report shows the staff statistic of PTB 9.3 International Cooperation in a gendered manner. **The goal of this report is to calculate whether the gender distribution by position can be attributed to chance and to calculate the unadjusted genderpay-gap.**\
Not the goal of this report is not to criticize the gender imbalance. This is just an exercise in data analysis and writing reports with R.

#### **Overview over Group 9.3**

When working with *PTB Group 9.3 International Cooperation* one is quickly under the impression: this is a very female Department, especially given the public image of the German federal administration as male dominated (Fig. 1).\

```{r Staff of all sections, echo=FALSE, fig.show='hold', out.width="50%"}
print(fig_bySection)
print(fig_position_gesamt)
```

But when examining the gender distribution by position (Fig. 2) a slight gender bias appears: there are next to none male Project Assistants, `r round(mean(personal$Sex[personal$Position == "PA"] == "M"), digits = 3) * 100`% to be exact, while males being between `r round(mean(personal$Sex[personal$Position == "PK"] == "M"), digits = 3) *100`% and `r round(mean(personal$Sex[personal$Position == "FB"] == "M"), digits = 3) *100`% of coworkers in the other positions at PTB 9.3.

#### $\chi^2$- Test for Homogeneity 

To further understand the gender distribution of Project Assistants we do a $\chi^2$-Test for homogeneity. To do a $\chi^2$- Test for homogeneity two tables are generated. The observed values table (Fig. 3) displays the distribution observed in PTB Group 9.3. The expected values table (Fig. 4) shows the distribution of male and female coworkers, if there was no impact of working in different positions. Not included are the 9.3 staff, only the staff within the sections will be included in the following analysis.

```{r Chi^2 Test, echo=FALSE, warning=FALSE}
subset = personal %>% filter(Section != "9.3")
chisq.sex.position = chisq.test(subset$Sex, subset$Position)
rm(subset)
d1 = kable(chisq.sex.position$observed, col.names = c("Sex", metadata$Long), caption = "Fig. 3 - Observed Values (PTB Group 9.3)", align = "c")
d2 = kable(chisq.sex.position$expected, col.names = c("Sex", metadata$Long), caption = "Fig. 4 - Expected Values (Calculated)", digits = 1, align = "c")
kables(list(d1, d2))
rm(d1, d2)
```

By calculating the difference between these two tables a $\chi^2$- Value is generated. This value is then tested for it\`s statistical significance.\
In this case the $\chi^2$- Value is `r round(chisq.sex.position$statistic, digits = 3)` indicating an effect. To gauge the significance of this value a P-Value is calculated. In this case the P-Value represents the probability of the observed values being distributed this was due to chance and not due to some effect. The P-Value of this $\chi^2$-Test is `r round(chisq.sex.position$p.value, digits = 3)` which equates to `r round(chisq.sex.position$p.value, digits = 3) * 100`%. `r round(chisq.sex.position$p.value, digits = 3) * 100`% is not a very high significance, but one that suggests a certain effect of gender on the choice of a position or vice versa.\
A $\chi^2$ value is dependent on the amount in cases and variables of the used data. Therefore to be able to gauge the strength of the effect, a standardization of $\chi^2$ is necessary. Commonly used is Cramer\`s V which normalizes $\chi^2$ to a range from 0 - 1, whereby 0 means no effect and 1 is the strongest possible effect.

```{r CramerV, echo=FALSE}
cramerV(chisq.sex.position$observed)
```

To conclude the $\chi^2$-Test: A medium sized effect between gender and position was calculated and this effect is significant to a high degree.

#### **Unadjusted Genderpay-Gap**

Now that an statistically significant effect between gender and position within *PTB 9.3 International Cooperation* is established let\`s assess it\`s impact. For this purpose we will calculate the undajusted genderpay-gap. This calculation will be a very rough estimate, because in Germany no individual public income statistic is available. But because the public sector operates with an openly available pay-table ([TVÖD Bund 2024](https://oeffentlicher-dienst.info/c/t/rechner/tvoed/bund?id=tvoed-bund-2024&matrix=1)) for different qualifications we can make an educated guess for each position.

```{r Genderpay-Gap, echo=FALSE}
metadata %>% select(! Short) %>% kable(col.names = c("Position", "Estimated minimal Group", 
                                                     "Estimated maximal Group", "Estimated Monthly min €", "Estimated Monthly max €"), 
                                       align = "c", caption = "Fig. 5 - Estimated Pay Groups by Position\n Based on TVÖD Bund 2024")
```
We will use the starting salary of each pay group respectively.  

With this information we can now calculate two estimates for the Unadjusted Genderpay-Gap of PTB 9.3. These estimates do not include the staff of PTB 9.3 itself. Only those working in the sections 9.31 - 9.35.
